---
title: "Classification methods for Brain-Computer Interface EEG data"
author: "Livio Finos, Marco Congedo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick tutorial for Rieman package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- \VignetteEngine{knitr::rmarkdown} -->

# Introduction

## Data Example
Available on 
<https://zenodo.org/record/1494240#.XARcSGhKg2x>

In the Brain Invaders P300 paradigm, a repetition is composed of 12 flashes, of which 2 include the Target symbol (Target flashes) and 10 do not (non-Target flash). For this experiment, in the Training phases the number of flashes is fixed (80 Target flashes and 400 non-Target flashes). In the Online phases the number of Target and non-Target still are in a ratio 1/5, however their number is variable because the Brain Invaders works with a fixed number of game levels, however the number of repetitions needed to destroy the target (hence to proceed to the next level) depends on the user’s performance. 

In any case, since the classes are unbalanced, an appropriate score must be used for quantifying the performance of classification methods (e.g., balanced accuracy, AUC methods, etc).


Data used: Subject 11

  - filename: '3.gdf'
    experimental_condition: nonadaptive
    type: training
  - filename: '4.gdf'
    experimental_condition: nonadaptive
    type: online    


```{r}
rm(list=ls())
# library(readr)
# train <- read_delim("C:/Users/livio/Downloads/sub11_s3_epocs.txt", delim = "\t", escape_double = FALSE, trim_ws = TRUE)
# train$X18=NULL
# # View(train)
# # table(train$X1)
# # plot(train$X1[(1:700)*1000],type="l")
# 
# 
# test <- read_delim("C:/Users/livio/Downloads/sub11_s4_epocs.txt", delim = "\t", escape_double = FALSE, trim_ws = TRUE)
# # View(test)
# # table(test$X1)
# test$X18=NULL
# 
# # check che i tempi coincidano:
# all(unique(test$X1)==unique(train$X1))
# 
# timepoints=unique(train$X1)
# ntimepoints=length(timepoints)
# 
# test=test[,-1,]
# dim(test)
# train=train[,-1]
# 
# test=as.matrix(test)
# train=as.matrix(train)
# 
# # rashape datasets as 3D array
# train2=array(train,c(ntimepoints,nrow(train)/ntimepoints,ncol(train)))
# train2=aperm(train2,c(1,3,2))
# test2=array(test,c(ntimepoints,nrow(test)/ntimepoints,ncol(test)))
# test2=aperm(test2,c(1,3,2))
# 
# save(test2,timepoints,ntimepoints,file="test2.Rdata")
# save(train2,timepoints,ntimepoints,file="train2.Rdata")

load(file="test2.Rdata")
load(file="train2.Rdata")

```

Eventi:

```{r}
library(readr)
eventi_s3 <- read_table2("C:/Users/livio/Dropbox (DPSS)/didattica/neuroQ/7_BCI/eventi_s3.csv")
# View(eventi_s3)
eventi_s4 <- read_table2("C:/Users/livio/Dropbox (DPSS)/didattica/neuroQ/7_BCI/eventi_s4.csv")
# View(eventi_s4)

#rename and reorder the lables
eventi_s3$type=factor(eventi_s3$type)
levels(eventi_s3$type)=c("Target","NonTarget")
eventi_s3$type=factor(eventi_s3$type,levels = c("NonTarget","Target"))

eventi_s4$type=factor(eventi_s4$type)
levels(eventi_s4$type)=c("Target","NonTarget")
eventi_s4$type=factor(eventi_s4$type,levels = c("NonTarget","Target"))

table(eventi_s3$type)
table(eventi_s4$type)

```

## Means

```{r}
Mtrain_target=apply(train2[,,eventi_s3$type=="Target"],c(1,2),mean)
Mtrain_nontarget=apply(train2[,,eventi_s3$type=="NonTarget"],c(1,2),mean)

dim(Mtrain_target)
dim(Mtrain_nontarget)

matplot(timepoints,Mtrain_target+matrix(1:16,byrow = TRUE,nrow(Mtrain_target),ncol(Mtrain_target)),type="l",lty=1,col=2)

matplot(timepoints,Mtrain_nontarget+matrix(1:16,byrow = TRUE,nrow(Mtrain_target),ncol(Mtrain_target)),type="l",lty=1,col=1)

matplot(timepoints,Mtrain_target-Mtrain_nontarget+matrix(1:16,byrow = TRUE,nrow(Mtrain_target),ncol(Mtrain_target)),type="l",lty=1,col=3)


```

## Select 0 to 650 ms

```{r}
selct_timepnts=which((timepoints>0)&(timepoints<650))

Mtrain_nontarget=Mtrain_nontarget[selct_timepnts,]
Mtrain_target=Mtrain_target[selct_timepnts,]

test2=test2[selct_timepnts,,]

```

## Euclidean distance

```{r}
dist_frobenius <- function(x,y){
  sqrt(sum((x-y)^2))
}

dist_frobenius(test2[,,1],Mtrain_nontarget)

# dist from nontarget - dist target.
# positive score means closer to target
diff_dist=sapply(1:dim(test2)[3],function(i)
  dist_frobenius(test2[,,i],Mtrain_nontarget)-dist_frobenius(test2[,,i],Mtrain_target))

plot(diff_dist,eventi_s4$type)

table(hat_target=diff_dist>=0,True_target=eventi_s4$type)

eventi_s4$estimate=diff_dist>=0
eventi_s4$estimate=factor(eventi_s4$estimate)
levels(eventi_s4$estimate)=levels(eventi_s4$type)

library(caret)
#Confusion Matrix
mat <- confusionMatrix(eventi_s4$estimate, eventi_s4$type, positive="Target")
#Confusion matrix
mat$table


# Sensitivity = TP/(TP+FN)
mat$byClass["Sensitivity"]

# Specificity = TN/(TN+FP)  
mat$byClass["Specificity"]


mat$overall["Accuracy"]

library(pROC)
roc.val <- roc(type~diff_dist, eventi_s4)
plot(roc.val, main="pROC package ROC plot") 
roc.val$auc


```

## Riemann

```{r}

# NON Target:
XNT=data.frame(cbind(test2[,,6],Mtrain_nontarget))

library(reshape2)
melted_cormat <- melt(cor(XNT))
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()


# Target:
XT=data.frame(cbind(test2[,,6],Mtrain_target))

library(reshape2)
melted_cormat <- melt(cor(XT))
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()


eventi_s4$type[6]


library(Riemann)
dist_Riemann(cor(test2[,,1]),cor(Mtrain_nontarget))
# dist from nontarget - dist target.
# positive score means closer to target
diff_dist_Riem=sapply(1:dim(test2)[3],function(i)
  dist_Riemann(cor(test2[,,i]),cor(Mtrain_nontarget))-dist_Riemann(cor(test2[,,i]),cor(Mtrain_target)))

plot(diff_dist_Riem,eventi_s4$type)

eventi_s4$estimate=diff_dist_Riem>=0
eventi_s4$estimate=factor(eventi_s4$estimate)
levels(eventi_s4$estimate)=levels(eventi_s4$type)

library(caret)
#Confusion Matrix
mat <- confusionMatrix(eventi_s4$estimate, eventi_s4$type, positive="Target")
#Confusion matrix
mat$table


# Sensitivity = TP/(TP+FN)
mat$byClass["Sensitivity"]

# Specificity = TN/(TN+FP)  
mat$byClass["Specificity"]


mat$overall["Accuracy"]

library(pROC)
roc.val <- roc(type~diff_dist, eventi_s4)
plot(roc.val, main="pROC package ROC plot") 
roc.val$auc

```

